Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2022-10-16 manuscript.tex
Code update with full OSF integration, added extra figure.
Updates to manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Different rules for binocular combination of luminance in cortical and subcortical pathways},
  pdfauthor={Federico G. Segala, Aurelio Bruno, Alex R. Wade \& Daniel H. Baker (+Myat? +Joel?)},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Different rules for binocular combination of luminance in cortical and subcortical pathways}
\author{Federico G. Segala, Aurelio Bruno, Alex R. Wade \& Daniel H. Baker (+Myat? +Joel?)}
\date{2022-10-04}

@@

@@

@@

Our experiment used different stimuli than the ones used by Anstis and Ho: in the binocular fusion experiment, their stimuli were not flickering and, in the flicker experiment, they were always shown in both eyes. In our experiment, in some conditions, the flicker was shown to only one eye. Moreover, they were looking at the temporal fusion of the flicker while we focussed on the binocular fusion of the flicker. Based on this and on the results from our first experiment, we expected to find a near linear summation of the responses.

@@

Thirty, twelve and ten participants were recruited for experiment 1, 2 and 3 respectively. All participants had normal or corrected to normal binocular vision. All participants gave written informed consent.

\hypertarget{apparatus-stimuli}{%
\subsection{Apparatus \& Stimuli}\label{apparatus-stimuli}}

The stimuli were two discs of flickering light with a diameter of 3.74 degrees, presented on a black background. The same stimuli were used for all three experiments. Four dark red lines were added around both discs to help with their fusion into one binocular disc (see insert in figure \ref{fig:pupildata}a for an example of the of the fused stimuli). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors and so did not suffer from internal reflections, and allowed participants to see only one fused disc. The use of a stereoscope allowed us to display the stimuli in three different ocular configurations: monocular, binocular, and dichoptic.

@@

EEG data were collected for experiments 1 and 2 sampled at 1 kHz using a 64-electrode ANT WaveGuard cap and the signals were recorded using the ASA software (ANT Neuro, Hengelo, NL, United States). Pupillometry data were collected only for experiment 1 using a binocular Pupil Core eye-tracker (Pupil Labs GmbH, Kassner et al. (2014)) running at 120 Hz and the signals were recorded with the Pupil Capture software (Pupil Labs GmbH, Kassner et al. (2014)).

@@

\hypertarget{experiment-1}{%
\subsubsection{Experiment 1}\label{experiment-1}}

The experiment was conducted in a windowless room, in which the only light source was the monitor. The participants sat at 99 cm from the monitor and their viewing distance through the stereoscope was 107 cm. The experiment was carried out in one session lasting 45 minutes in total, divided in three blocks of 15 minutes each. In each block, there were 60 trials in total lasting 15 seconds each (12s of stimulus presentation, with an interstimulus interval of 3s). The participants had no task and were asked to look at the fixation crosses in the middle of the two discs while trying to minimise their blinking during the presentation period. The discs flickered in counterphase at 2 Hz, but in some conditions (cross-frequency conditions) one of the two discs flickered at 1.6 Hz. They were set at five temporal contrast levels relative to the mean luminance: 6, 12, 24, 48 and 96 \%.

As already stated, the mirror stereoscope allowed the use of three different ocular configurations. In the monocular configuration, one disc was set to flicker either at 2 Hz or 1.6 Hz for each contrast level, while the other disc remained stable at a constant mean luminance. In the binocular configuration, both discs flickered at the same contrast level with both discs flickering at 2 Hz in the same-frequency condition, while, in the cross-frequency condition, one disc flickered at 2 Hz and the other at 1.6 Hz. Finally, in the dichoptic configuration, one disc flickered across all contrast levels and the other flickered at a fixed contrast level that was set at 48 \%. Similar to the binocular configuration, the dichoptic configuration was divided in same-frequency and cross-frequency conditions.

\hypertarget{experiment-2}{%
\subsubsection{Experiment 2}\label{experiment-2}}

The conditions were similar to experiment 1 (windowless room and viewing distance of 107 cm). Like in experiment 1, participants had no specific task other than looking at the fixation crosses and minimising their blinking. Unlike the first experiment, only one contrast level was used (96 \%) and the discs were set to flicker at five different frequencies (2, 4, 8, 16 and 30 Hz). Only two ocular configurations, monocular and binocular, were used, with the latter having both discs flickering at the same frequency.

The experiment was carried out in one session lasting 25 minutes in total, divided in five blocks of 5 minutes each. In each block, there were 20 trials in total lasting 15 seconds each (12s of stimulus presentation and 3s of interstimulus interval).

\hypertarget{experiment-3}{%
\subsubsection{Experiment 3}\label{experiment-3}}

The experiment was conducted in a room with a blacked-out window. The sitting and viewing distances of the participants were similar to the two previous experiments. As this was a psychophysical task, no EEG or pupillometry data were collected for this experiment.

A two-interval forced-choice staircase procedure was used to collect data. In one interval, participants were presented with a standard fused disc that flickered at a set contrast level (either 24 or 48 \%), which was selected by the experimenter at the beginning of each block. In a separate interval, a target fused disc flickering at varying contrast levels was displayed. The contrast level of the target was controlled by a 1-up, 1-down staircase moving in logarithmic (dB) steps of contrast. The ratio of flicker amplitudes in the left and right eyes was varied across blocks and was set to be 0, 0.25, 0.5, 0.75 or 1. The standard and the target fused discs were displayed for 1 second each, with an interstimulus interval of 0.5 seconds. After the discs appeared on screen, the participants had to indicate which interval they perceived as having the more intense flicker. The intervals were randomly ordered, and all discs flickered at a frequency of 2 Hz.

Due to its long duration (162 minutes in total), the participants completed the experiment across multiple sessions divided at their own discretion. The experiment was divided in 54 blocks (3 repetitions \(\times\) 2 standard contrasts \(\times\) 9 target ratios), which lasted on average 3 minutes each (depending on the response speed of the participant). In each block, there were a total of 50 trials. We pooled the data from all repetitions of a given condition and fitted a cumulative normal psychometric function to estimate the point of subjective equality at the 50 \% point.

@@

EEG data were converted from the ANT-EEProbe format to a compressed csv text file using a custom Matlab script and components of the EEGlab toolbox (Delorme and Makeig, 2004). The data for each participant were then loaded into R for analysis where a ten-second waveform for each trial at each electrode was extracted. Fourier transform was calculated for each waveform, and the spectrum stored in a matrix. All repetitions of each condition were then averaged for each electrode. They were then averaged across the four occipital electrodes (POz, Oz, O1, O2), to obtain individual results. Finally, these were averaged across all participants to obtain the final results.

A similar process was adopted for the analysis of the pupillometry data. The data were converted from the mp4 format to a compressed csv text file using the Pupil Player software (Kassner et al., 2014). The individual data were then loaded into R for analysis where a twelve-second waveform for each trial in each eye was extracted. Fourier transform was calculated for each waveform, and the spectrum stored in a matrix. All repetitions of each condition were then averaged for each eye. They were then averaged across the two eyes, to obtain individual results. Finally, these were averaged across all participants to obtain the final results.

@@

\hypertarget{experiment-1-1}{%
\subsection{Experiment 1}\label{experiment-1-1}}

The results from the first experiment are shown in figure \ref{fig:pupildata} and figure \ref{fig:EEGdata}. Each curve is showing the amplitude of the signal for the three different configurations (monocular, binocular and dichoptic).

Figure \ref{fig:pupildata} is showing the results for the pupillometry. Figure \ref{fig:pupildata}a is showing the group average waveform for binocular presentation. Figure \ref{fig:pupildata}b is showing the average Fourier spectrum, where we can observe a spike at 2 Hz, which shows a very clear, non-noisy response to the flickering stimuli. Figures \ref{fig:pupildata}c-e show the contrast response functions (CRFs) at 2 and 1.6 Hz. In the same-frequency condition at 2 Hz (fig.~\ref{fig:pupildata}c), the amplitude of the binocular configuration is bigger than the amplitude of the monocular configuration. In general, we observe that the binocular to monocular ratio is roughly \(\sqrt2\) at each contrast level (fig.~).
In the cross-frequency condtion at 2 Hz (fig.~\ref{fig:pupildata}d), no summation is observed since the amplitudes of the binocular and dichoptic conditions are smaller than the amplitude of the monocular condition. Interestingly, we observe a relatively small cross-frequency facilitation at 1.6 Hz (fig.~\ref{fig:pupildata}e), with the binocular amplitude being bigger than the monocular amplitude, especially at the lower contrast levels. This facilitation is then lost at the highest contrast level, where the amplitude of the monocular signal is equal to the amplitude of the binocular signal. This pattern of results in the cross-frequency conditions suggests the presence of interocular suppression at the level of the pupils. Moreover, the low summation observed in the same-frequency condition suggests that theprocessing of temporal luminance modulations happens in a non-linear way at the level of the pupils.

Figure \ref{fig:EEGdata} shows the results for EEG. Figure \ref{fig:EEGdata}a is showing the group average waveform for binocular presentation. Figure \ref{fig:EEGdata}b is showing the average Fourier spectrum, where we can observe two spikes at 2 and 4 Hz, which show a very clear, non-noisy response to the flickering stimuli at the first and second harmonics. Figures \ref{fig:EEGdata}c-e show the CRFs at 2 and 1.6 Hz at the first harmonic, and figures \ref{fig:EEGdata}f-h show the CRFs at the second harmonic.
What is immediately apparent is that, in the same-frequency condition, the binocular responses at the first (fig.~\ref{fig:EEGdata}c) and second (fig.~\ref{fig:EEGdata}f) harmonics are showing very high summation of binocular signals compared to the pupillometry. The amplitudes of the binocular responses are twice, and even more than twice, the amplitudes of the monocular responses. This is especially clear at the highest contrasts. This is clearly displayed in figure, which is showing the binocular to monocular ratios for the EEG at both harmonics and the pupillometry. For the cross-frequency condition at 2 Hz and 4 Hz (fig.~\ref{fig:EEGdata}d and \ref{fig:EEGdata}g), there is almost no summation, with the amplitudes of the binocular and dichoptic configurations being equal, and even smaller, than the amplitudes of the monocular signals. This pattern is also observable in the cross-frequency condition at 1.6 (fig.~\ref{fig:EEGdata}e) and 3.2 Hz (fig.~\ref{fig:EEGdata}h). This pattern of results seems to suggest that processing of temporal luminance modulations happens in a linear way in the visual cortex. Moreover, the lower amplitudes for the binocular and the dichoptic configurations in the cross-frequency conditions seem to indicate the presence of interocular suppression in the visual cortex.

@@

\caption{Summary of pupillometry results for N=30 participants. Panel (a) shows a group average waveform for binocular presentation (low pass filtered at 5Hz), with the driving signal plotted at the foot. Panel (b) shows the average Fourier spectrum. Panels (c,d) show contrast response functions at 2Hz for different conditions. Panel (e) shows contrast response functions at 1.6Hz for three conditions. Shaded regions and error bars indicate bootstrapped standard errors.}\label{fig:pupildata}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Different rules for binocular combination of luminance flicker in cortical and subcortical pathways},
  pdfauthor={Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Different rules for binocular combination of luminance flicker in cortical and subcortical pathways}
\author{Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker}
\date{2022-10-16}

@@

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

OK, some pointers for the introduction:

I think we need to draw a distinction between binocular summation at threshold (on which there is a big literature) and binocular facilitation above threshold. Much of the introduction is focussed on threshold measures, but actually I think our focus here is more about suprathreshold conditions where we usually see little facilitation. I think it will help the reader if we use the term summation only to refer to threshold phenomena, and facilitation in the more general case (I've tried to do this elsewhere in the paper).

The key points should be:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  For grating stimuli, there are summation effects at threshold, but these are lost at high contrasts.
\item
  Binocular combination happens in multiple pathways, including the pupil pathway. (This is a good place to mention the Quaia paper).
\item
  We don't know much about the pupil response, though it is clearly binocular because of the consensual response (convergence and divergence seem to be other ways of saying the same thing, so not sure they are worth mentioning).
\item
  Actually there's not much work on temporal binocular combination even in the primary visual pathway.
\item
  We have developed a novel paradigm that allows us to probe both cortical and subcortical pathways simultaneously for flickering light.
\end{enumerate}

I wouldn't say too much about models, or about the matching experiment, at this stage.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

Several models have been developed and proposed to describe binocular combination and contrast matching results for spatial contrast. Among these, the models proposed by Ding and Sperling (2006) and Meese and colleagues (2006) are widely accepted. These models incorporate a dynamic contrast gain control, which allows to develop a model that combines binocular summation and interocular suppression The first model is a multistage model that consists of two pairs of gain control mechanisms that inhibit each other (Ding and Sperling, 2006). In the earlier stage, the right and left eye channels exert gain control on the other channel. Then, they exert gain control on the other channel's gain control. Lastly, the outputs from each channel are summed binocularly to determine the binocular signal's magnitude. The second model is similar as it also incorporates gain control mechanisms (Meese et al., 2006). In the first stage, it is applied to the left and right eye channels with suppression from the other eye. Then the signal from the two eyes is summed binocularly and gain control is applied a second time. This second model in particular is interesting and relevant as it was designed to also explain data from matching experiments.

@@

Our experiment used different stimuli than the ones used by Anstis and Ho (1998): in the binocular fusion experiment, their stimuli were not flickering and, in the flicker experiment, they were always shown in both eyes. In our experiment, in some conditions, the flicker was shown to only one eye. Moreover, they were looking at the temporal fusion of the flicker while we focussed on the binocular fusion of the flicker. Based on this and on the results from our first experiment, we expected to find a near linear summation of the responses.

@@

Thirty, twelve and ten participants were recruited for Experiments 1, 2 and 3 respectively. All participants had normal or corrected to normal binocular vision, and gave written informed consent.

\hypertarget{apparatus-stimuli}{%
\subsection{Apparatus \& stimuli}\label{apparatus-stimuli}}

The stimuli were two discs of flickering light with a diameter of 3.74 degrees, presented on a black background. The same stimuli were used for all three experiments. Four dark red lines were added around both discs to help with their fusion into one binocular disc (see insert in Figure \ref{fig:pupildata}b for an example of the fused stimulus). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors to avoid internal reflections, and meant that participants saw a single fused disc. The use of a stereoscope allowed us to modulate the stimuli in three different ocular configurations: monocular, binocular, and dichoptic. Note that during monocular presentation of flicker, the unstimulated eye still saw the static (non-flickering) disc of mean luminance.

@@

EEG data were collected for Experiments 1 and 2 using a 64-electrode ANT WaveGuard cap and the signals were recorded at 1 kHz using the ASA software (ANT Neuro, Netherlands). Pupillometry data were collected for Experiment 1 using a binocular Pupil Core eye-tracker (Pupil Labs GmbH, Berlin, Germany; Kassner et al. (2014)) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

\hypertarget{experiment-1-simultaneous-eeg-and-pupillometry}{%
\subsubsection{Experiment 1: simultaneous EEG and pupillometry}\label{experiment-1-simultaneous-eeg-and-pupillometry}}

The experiment was conducted in a windowless room, in which the only light source was the monitor. The participants sat at 99 cm from the monitor and the total optical viewing distance (through the stereoscope) was 107 cm. The experiment was carried out in one session lasting 45 minutes in total, divided in three blocks of 15 minutes each. In each block, there were 60 trials in total lasting 15 seconds each (12s of stimulus presentation, with an interstimulus interval of 3s). The participants were given no specific task instructions, but were asked to look at the fixation crosses in the middle of the two discs while trying to minimise their blinking during the presentation period.
We included six distinct ocular conditions, each at five temporal target contrast levels (combined factorially) relative to the mean luminance: 6, 12, 24, 48 and 96\%. In the first three conditions, the discs flickered at 2 Hz, in either a monocular, binocular, or dichoptic arrangement. In the dichoptic condition the non-target eye saw a fixed luminance contrast of 48\%. In the remaining three conditions (termed the cross-frequency conditions) one eye's disc flickered at 1.6Hz, and the other eye's disc flickered at 2Hz. We also tested monocular (one eye sees 1.6Hz flicker, the other sees mean luminance), binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 2Hz, mask contrast of 48\% at 1.6Hz in the other eye) arrangements. The rationale for flickering both eyes at 2Hz is that we can then measure summation behaviour between the eyes in the pupil and EEG response at 2Hz. The rationale for flickering the eyes at different frequencies is that this permits measurement of suppression between the eyes (i.e.~the reduction in the 2Hz response when a 1.6Hz mask component is added to the other eye). We counterbalanced presentation of the target stimulus across the left and right eyes.

\hypertarget{experiment-2-eeg-responses-across-temporal-frequency}{%
\subsubsection{Experiment 2: EEG responses across temporal frequency}\label{experiment-2-eeg-responses-across-temporal-frequency}}

This experiment used the same equipment set up as Experiment 1, except that the eye tracker was not used. Unlike the first experiment, only one contrast level was used (96\%) and the discs were set to flicker at five different frequencies (2, 4, 8, 16 and 30 Hz). Only two ocular configurations, monocular and binocular, were used, with the latter having both discs flickering at the same frequency. The experiment was carried out in one session lasting 25 minutes in total, divided into five blocks of 5 minutes each. In each block, there were 20 trials in total with the same timing as for Experiment 1.

\hypertarget{experiment-3-temporal-contrast-matching}{%
\subsubsection{Experiment 3: temporal contrast matching}\label{experiment-3-temporal-contrast-matching}}

The experiment was conducted in a darkened room with a blacked-out window. The display equipment (monitor and stereoscope) were the same as for the two previous experiments, but no EEG or pupillometry data were collected. A two-interval contrast matching procedure was used to collect data. In one interval, participants were presented with a standard fused disc that flickered at a set contrast level (either 24 or 48\%), which was selected by the experimenter at the beginning of each block. In the other interval, a target disc flickering at varying contrast levels was displayed. The contrast level of the target was controlled by a 1-up, 1-down staircase moving in logarithmic (dB) steps of contrast. The ratio of flicker amplitudes in the left and right eyes was varied across blocks and was set to be 0, 0.25, 0.5, 0.75 or 1 (9 distinct conditions). The standard and target discs were displayed for 1 second each, with an interstimulus interval of 0.5 seconds. After the discs appeared on screen, the participants had to indicate which interval they perceived as having the more intense flicker. The intervals were randomly ordered, and all discs flickered at a frequency of 2 Hz (two cycles in sine phase).

Due to its long duration (approximately 3 hours in total), the participants completed the experiment across multiple sessions initiated at their own convenience The experiment was divided into 54 blocks (3 repetitions \(\times\) 2 standard contrasts \(\times\) 9 target ratios), which lasted on average 3 minutes each, depending on the response speed of the participant. In each block, there were a total of 50 trials. No auditory feedback was given for this subjective task.

@@

EEG data were converted from the ANT-EEProbe format to a compressed csv text file using a custom Matlab script and components of the EEGlab toolbox (Delorme and Makeig, 2004). The data for each participant were then loaded into R for analysis where a ten-second waveform for each trial at each electrode was extracted (omitting the first two seconds). The fast Fourier transform was calculated for each waveform, and the spectrum stored in a matrix. All repetitions of each condition were then averaged for each electrode. They were then averaged across four occipital electrodes (\emph{POz}, \emph{Oz}, \emph{O1}, \emph{O2}), to obtain individual results. Finally, these were averaged across participants to obtain the group results. All averaging retained the phase information (i.e.~coherent averaging), and at each stage we excluded data points with a Mahalanobis distance exceeding \(D\) = 3 from the complex-valued mean (see Baker, 2021). For statistical comparisons of complex-valued data, we use the \(ANOVA^2_{circ}\) statistic described by Baker (2021). This is a multivariate extension of ANOVA that assumes equal variance of the real and imaginary Fourier components, or equivalently, an extension of the \(T^2_{circ}\) statistic of Victor and Mast (1991) that can compare more than two conditions.

A similar analysis pipeline was adopted for the pupillometry data. The data were converted from mp4 videos to a csv text file using the Pupil Player software (Kassner et al., 2014), which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where again a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The fast Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. Again, we used coherent averaging, and excluded outlying data points in the same way as for the EEG data. Note that previous pupillometry studies using luminance flicker have tended to fit sine-waves to the data, rather than using Fourier analysis (e.g. Spitschan et al., 2014). Our novel approach is more robust to noise at other frequencies, and makes the pupillometry analysis consistent with standard practice in steady-state EEG analysis (e.g. Figueira et al., 2022)

To analyse the matching data, we pooled the trial responses across all repetitions of a given condition for each participant. We then fitted a cumulative normal psychometric function to estimate the point of subjective equality at the 50\% level. Thresholds were averaged across participants in logarithmic (dB) units. For all experiments, we used a bootstrapping procedure with 1000 iterations to estimate standard errors across participants. All analysis and figure construction was conducted using a single R-script, available online, making this study fully computationally reproducible.

\hypertarget{computational-model-and-parameter-estimation}{%
\subsection{Computational model and parameter estimation}\label{computational-model-and-parameter-estimation}}

To describe our data, we chose a model of binocular contrast gain control with the same general form as the first stage of the model proposed by Meese et al. (2006). The second gain control stage was omitted (consistent with Baker and Wade, 2017) to simplify the model and reduce the number of free parameters. The response of the left eye's channel is given by:

\begin{equation}
\label{eq:respL}
Resp_L = \frac{L^2}{Z + L + wR},
\end{equation}

with an equivalent expression for the right eye:

\begin{equation}
\label{eq:respR}
Resp_R = \frac{R^2}{Z + R + wL}.
\end{equation}

In both equations, \emph{L} and \emph{R} are the contrast signals from the left and right eyes, \emph{Z} is a saturation constant that shifts the contrast-response function laterally, and \emph{w} is the weight of suppression from the other eye.

The responses from the two eyes are then summed binocularly:

\begin{equation}
\label{eq:respB}
Resp_B = R_{max}(Resp_L + Resp_R) + n,
\end{equation}

where \emph{n} is a noise parameter, and \(R_{max}\) scales the overall response amplitude. The \(R_{max}\) parameter was omitted when modelling the contrast matching data, as it had no effect in this paradigm.

Despite being derived from the model proposed by Meese et al. (2006), the simplifications applied to this architecture make it very similar to other models (e.g. Ding and Sperling, 2006; Legge, 1984; Schrödinger, 1926). In particular we fixed the numerator exponent at 2 in our model, because otherwise this value tends to trade off with the weight of interocular suppression (see Baker et al., 2012; Kingdom and Libenson, 2015). Our key parameter of interest is the weight of interocular suppression. Large values around \(w\) = 1 result in a very small or nonexistent binocular advantage at suprathreshold contrasts, consistent with previous work using grating stimuli (Baker and Wade, 2017). Low values around \(w\) = 0 produce substantial, near-linear binocular facilitation (Baker et al., 2020).

We implemented the model within a Bayesian framework using the Stan software (Carpenter et al., 2017). This allowed us to estimate group-level posterior parameter distributions for the weight of interocular suppression, \(w\), and the other free model parameters \(R_{max}\), \(Z\) and \(n\). The prior distributions for all parameters were Gaussian, with means and standard deviations of 1 and 0.5 for \(w\) and \(R_{max}\), and 5 and 2 for \(Z\) and \(n\). We sampled from a Student's t-distribution for the amplitudes in the pupillometry and EEG experiments, and from a Bernoulli distribution for the single trial matching data. The models were fit using the individual data across all participants, independently for each data set. Because we used coherent averaging across participants, the group average amplitudes are shifted vertically relative to the model predictions, which are based on hierarchical fits to the individual participant amplitudes (put another way, the model does not implement coherent averaging). However it is clear that the character of the model in all cases gives a good representation of the data. We sampled over a million steps for each data set, using a computer cluster, and retained 10\% of samples for plotting.

\hypertarget{preregistration-data-and-code-availability}{%
\subsection{Preregistration, data and code availability}\label{preregistration-data-and-code-availability}}

We initially preregistered our main hypotheses and analysis intentions for the first experiment. We then conducted a pilot study with N=12 participants, before making some minor changes to the stimulus (we added dim red lines to aid binocular fusion). We then ran the main experiment, followed by two additional experiments that were not preregistered. The preregistration document, raw data files, and experimental and analysis code, are available on the project repository: \url{https://osf.io/tbema/}.

@@

\hypertarget{experiment-1}{%
\subsection{Experiment 1}\label{experiment-1}}

The pupillometry results are summarised in Figure \ref{fig:pupildata}. The group average waveform for binocular presentation is shown in Figure \ref{fig:pupildata}a. There is a substantial pupil constriction at stimulus onset, followed by visible oscillations at the flicker frequency (2Hz, see waveform at foot). The average Fourier spectrum is displayed in Figure \ref{fig:pupildata}b, and shows a substantial spike at 2 Hz, but no evidence of a second harmonic response at 4Hz. These results demonstrate that our paradigm can evoke measurable steady-state pupil responses at 2Hz.

@@

\caption{Summary of pupillometry results for N=30 participants. Panel (a) shows a group average waveform for binocular presentation (low pass filtered at 5Hz), with the driving signal plotted at the foot. Panel (b) shows the average Fourier spectrum, with an inset image illustrating the stimulus appearance (upper right). Panels (c,d) show contrast response functions at 2Hz for different conditions. Panel (e) shows contrast response functions at 1.6Hz for three conditions. Shaded regions and error bars indicate bootstrapped standard errors.}\label{fig:pupildata}
\end{figure}

Figure \ref{fig:pupildata}c shows contrast response functions in response to stimuli flickering only at 2Hz. Response amplitudes increased monotonically with target contrast, confirming that our paradigm is suitable for measuring contrast-dependent differences in response (to our knowledge this is the first time this has been demonstrated). The amplitude of the binocular condition (blue squares) is consistently greater than that of the monocular condition (red circles) across all target contrasts. A \(2\times5\) repeated measures \(ANOVA^2_{circ}\) (Baker, 2021) comparing these conditions revealed a significant main effect of target contrast (F(8,580) = 16.79, \(p\) \textless{} 0.001), a significant effect of condition (F(2,580) = 11.04, \(p\) \textless{} 0.001), and a significant interaction (F(8,580) = 56.25, \(p\) \textless{} 0.001). The dichoptic condition begins at a much higher amplitude, owing to binocular combination of the target and high (48\%) contrast mask, and then increases slightly with increasing target contrast (main effect of target contrast: F(8,232) = 3.03, \(p\) \textless{} 0.003).
